Process-based phenotypes: an application to the ontology of cell physiology

### 2011-11-26 trunk/paper/drinking.tex

In the domain of phenotypes, multiple ontologies have been
developed. In particular, ontologies to characterize mammalian
\cite{Smith2004}, human \cite{Robinson2008}, yeast \cite{ypo} and worm
\cite{wpo} phenotypes are now available, with several more ontologies
being under development.

structures, qualities

>>>>

In the domain of phenotypes, multiple ontologies have been
developed. In particular, ontologies to characterize mammalian
\cite{Smith2004}, human \cite{Robinson2008}, yeast \cite{ypo} and worm
\cite{wpo} phenotypes are now available, while several more phenotype
ontologies are under development. To benefit from automated reasoning,
integrate phenotypes across species and reuse the content of anatomy
and process ontologies, classes in phenotype ontologies were defined
using the framework of the Phenotypic Attribute and Trait (PATO)
ontology \cite{Gkoutos2005}. According to the PATO framework, a
phenotype can be decomposed, using an Entity-Quality model, into an
affected entity and a quality that characterizes {\em how} the entity
is affected \cite{Gkoutos2005}. Such decompositions have been created
for several widely used phenotype ontologies \cite{Mungall2010,
  Gkoutos2009b, obml2011h1}, and are being applied together with
methods for reusing knowledge contained in anatomy ontologies
\cite{Mungall2010, Hoehndorf2010phene}.

While the PATO framework is now successfully being applied to the
integration of phenotypes in which
